Reproducible Research Course Project 1
========================================
First we gonna loading and preprocessing the data

```{r}
library(ggplot2)
library(readr)
activity <- read_csv("activity.csv")

activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
```
Then we gonna answer some questions about the data 

##1. What is mean total number of steps taken per day?

```{r}
steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(steps) <- c("date", "steps")
qplot(steps$steps,main = "Total number of steps taken per day", xlab = "Steps")

mean(steps$steps)
median(steps$steps)
```
So, how we can see the mean of the total steps taken per day is 9354.23 

##2. What is the average daily activity pattern?
```{r}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
qplot(average_daily_activity$interval, average_daily_activity$mean, geom = "line",ylab="Average number of steps", xlab="intervals")

average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```


Here we can see the pattern of the average daily number of steps and the max numer of steps ofthe intervals of  5 minutes.  


##3. Imputing missing values

There are some days where there are missing values. We are gonna to calculate the total of this days.
```{r}
sum(is.na(activity$steps))
```
Then we construct a dataset with that values

```{r}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

qplot(total_steps_imputed$daily_steps,main = "Total number of steps taken each day", xlab = "Steps")

mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)


```

The difference betwen the first values an the newest is very sgnificant for the mean, but for the median is not very big because its sturdiness. 

##4. Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "sabado" | weekdays(x) =="domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

Making a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(.~datetype, ncol = 1, nrow=2)
print(plot)
```

